---
title: "Module 1.1 Introduction to Discrete Probability"
output: html_notebook
---
#Introduction to Discrete Probability
##Discrete Probability Video
More useful in games of chance.
Data science often uses continuious or numeric probability. 
Pr(A) - Notation used to show probability of an event
Event - Anything that can happen.
##13.1 Discrete Probability Textbook
##13.1.1 Relative Frequency
Example:  We have 2 red beads and 3 blue beads in a jar.  What is the probability of drawing red? Of drawing blue?
The event would be the specific draw:
Draw Red = 2/5 or .4
Draw Blue = 3/5 ot .6
##13.1.2 Notation
Pr(A) - Probability of specific eventA
To accurately collect probability information we should also express a specific event in some math term.

#Monte Carlo Simulation
Sample function in R will generate random draws from our data set. 
  Note - Default will be without replacement.  
    Sample with replace argument will replace the used values
Monte Carlo Simulation - How many times do we need to repeat a random sample before it is a karge enough repetition to be accurate
Replicate function - sets how many times we want to repeat an event.
##Bead Example
```{r}
# Create objects
#create our 'urn' with 2 red and 3 blue beads
beads <-rep(c("red" , "blue") , times = c(2 , 3))
# Display the 'bead' urn
beads
# Now lets sample 1 bead at random from beads
sample(beads, 1)
# Now lets do lots of draws
#set how many times
trials <- 10000
events <- replicate(trials, sample(beads , 1))
#create a table to store the events
tblEvents <- table(events)
#Show the table we created
tblEvents
#Show the properties of the table
prop.table(tblEvents)

```
```{r}
#Create the Scrubs dataset
# The different types of scrubs and the number of each pair
myScrubs <-rep(c("Black M&M" , "Black Nice" , "Royal Blue" , "Naby Blue") , times = c(1 , 2 , 1 , 2))
myScrubs #This will show all of my scrubs
#How many days do I work a year
workDays <- 365 #Number of days I work that year
# Define the event as me getting dressed.
##Replicate the number of work days we defined , sample from the my scrubs table, 1 time
getDressed <-replicate(workDays , sample(myScrubs , 1))
# create a table with the results of getDressed
tblScrubsWorn <- table(getDressed)
tblScrubsWorn
prop.table(tblScrubsWorn)
```
<<<<<<< HEAD
##Set the Random Seed
In some of these examples we are going to use random number generators.  This can mean that the results we get may not match the examples
set.seed will allow us to force the system to use a specific random number system.
Often people will use the date
```{r}
set.seed(1)
set.seed(1 , sample.kind = "Rounding")
```
This message is normal.  It is telling you that the non-standard method is being used. 

##Mean function
The mean function can often be used to detirmine probability. 
```{r}
beads <-rep(c("red" , "blue"), times = c(2 , 3))
beads
mean(beads == "blue")
mean(beads== "red")

```
When the mean function is used the variable you use with the function is compared against the array. 
A FALSE is given a 0 while a TRUE is given a 1
It then calculates the proportion of TRUE and returns the result. 
##Probability Distributions
When everything is discrete and in categories it is relatively straightforward to use a probability distribution.  It is the percentage of each category compared to the whole. 
Lets create a sample for politifcal party voting. 
```{r}
partyMakeup <-rep(c("Republican" , "Democrat" , "Undecided" , "Green") , times = c(44 , 44 , 10 , 2))
partyMakeup
population = 10000
populationbig <- 1000000
votes <- replicate(population, sample(partyMakeup , 1))
tblVotes <- table(votes)
#tblVotes
prop.table(tblVotes)

votesbig <-replicate(populationbig , sample(partyMakeup , 1))
tblVotesBig <-table(votesbig)
prop.table(tblVotesBig)

```
##Independence
Data independence is when the results of one event do not interfeer with the next.  
If we replace the bead we drew in the previous example each draw the outcome is independent.  
If we do not replace the bead, it will change the probability for the next draw because there will be one less total beads, and one less of the color you drew.  
Conditional Probabilities - When the probability of the second event is changed by the results of the first event
Notation:
  Pr(A|B) = Pr(A) x Pr(A|B)
  
Blackjack is a good example.
What is the probability of the first card being an Ace (A) and the second card being worth 10points (10, Jack, Queen, King)
Since the cards are not replaced, the events are not independent.  
Pr(AceFirst) = 4/52
Pr(10-J-Q-K) = 16/51
##Independence Textbook
It is critical that when we calculate the probabilities we do not assume dependence or independence.  This can lead to errors in our results. 
Lets retest the beads example if we do not replace
```{r}
beadSample <-3
x <- sample(beads, beadSample)
x
```
In this case the probability of the second/third/fourth/etc event is altered by the previous events.

##13.4 Conditional Probabilities 
When events are not independent then conditional probabilities are useful. 
When you see the | read as 

##13.5 Addition and Multiplication Rules
##13.5.1 Multiplication Rule
If we want the probability of 2 events both happening we use the multiplication rule.
This assumes the events are not independent.
Pr(A and B) = Pr(A) x Pr(B|A)

##13.5.2 Multiplication rule Under Independence
Pr(A and B and C) = Pr(A) x Pr(B) x Pr(C)
=======
Lets try to create a dice roll game.
>>>>>>> bb58af1bfa98c0d6b371747db768064d1bee6b65

##Combinations and Permutations (Video)
