---
title: "Module 1.1 Introduction to Discrete Probability"
output: html_notebook
---
#Introduction to Discrete Probability
##Discrete Probability Video
More useful in games of chance.
Data science often uses continuious or numeric probability. 
Pr(A) - Notation used to show probability of an event
Event - Anything that can happen.
##13.1 Discrete Probability Textbook
##13.1.1 Relative Frequency
Example:  We have 2 red beads and 3 blue beads in a jar.  What is the probability of drawing red? Of drawing blue?
The event would be the specific draw:
Draw Red = 2/5 or .4
Draw Blue = 3/5 ot .6
##13.1.2 Notation
Pr(A) - Probability of specific eventA
To accurately collect probability information we should also express a specific event in some math term.

#Monte Carlo Simulation
Sample function in R will generate random draws from our data set. 
  Note - Default will be without replacement.  
    Sample with replace argument will replace the used values
Monte Carlo Simulation - How many times do we need to repeat a random sample before it is a karge enough repetition to be accurate
Replicate function - sets how many times we want to repeat an event.
##Bead Example
```{r}
# Create objects
#create our 'urn' with 2 red and 3 blue beads
beads <-rep(c("red" , "blue") , times = c(2 , 3))
# Display the 'bead' urn
beads
# Now lets sample 1 bead at random from beads
sample(beads, 1)
# Now lets do lots of draws
#set how many times
trials <- 10000
events <- replicate(trials, sample(beads , 1))
#create a table to store the events
tblEvents <- table(events)
#Show the table we created
tblEvents
#Show the properties of the table
prop.table(tblEvents)

```
```{r}
#Create the Scrubs dataset
# The different types of scrubs and the number of each pair
myScrubs <-rep(c("Black M&M" , "Black Nice" , "Royal Blue" , "Naby Blue") , times = c(1 , 2 , 1 , 2))
myScrubs #This will show all of my scrubs
#How many days do I work a year
workDays <- 365 #Number of days I work that year
# Define the event as me getting dressed.
##Replicate the number of work days we defined , sample from the my scrubs table, 1 time
getDressed <-replicate(workDays , sample(myScrubs , 1))
# create a table with the results of getDressed
tblScrubsWorn <- table(getDressed)
tblScrubsWorn
prop.table(tblScrubsWorn)
```
Lets try to create a dice roll game.

