---
title: "Module 1.2 Combinations and Permutations"
output: html_notebook
---
#1.2 Combinations and Permutations
Combinations and Permutations are for more complex events and multiple events. 
paste()  - joins two strings together
expand.grid()  - gives the combinations of 2 vectors or lists
permutation(n,r) - lists the different ways that r items can be selected from a given set WHEN ORDER MATTERS!
combinations(n,r) - lists the different combinations when oth order does not matter. 

The gtools library contains these tools. 
Examples of paste and expand.grid
```{r}
library(gtools)
number <- "3"
suit <- "Hearts"
pasted <- paste(number , suit)
pasted

```
This will join the two vectors together. 

Joining vectors elemnt-wise with paste
```{r}
paste(letters[1:5] , as.character(1:5))
```
This generates a vector 1-5.  It then generates the first 5 letters.
It then pastes the two together ordered by elements. 

Expand Grid is used to combine multiple vectors. 
```{r}
clothes <- expand.grid(pants = c("blue" , "black") , shirts = c("white" , "grey" , "plat"))

clothes


```
```{r}
classes <-c("cleric" , "fighter" , "mage" , "rogue")
race <- c("human" , "dwarf" , "elf")
partyMembers <- expand.grid(race , classes)
partyMembers

```

###Permutations and Combinations
```{r}
library(gtools)
combinations(5,2)
#permutations(5,2) #Choose 2 in order from 5
all_phone_numbers <-permutations(10 , 7, v=0:9)
n <- nrow(all_phone_numbers)
index <- sample(n , 5)
all_phone_numbers[index,]


                
```
#Deck of Cards
This will generate a deck of cards that we can use for further analysis. 
```{r}
suits <- c("Diamonds" , "Hearts" , "Clubs" , "Spades")
numbers <- c("Ace" , "King" , "Queen" , "Jack" , "10" , "9" , "8" , "7" , "6", "5" , "4" , "3" , "2")
deck <-expand.grid(number = numbers , suit = suits)
deck <- paste(deck$number , deck$suit)
deck #print the deck to validate
kings <-paste("King" , suits)
mean(deck %in% kings)


```
What is the probability of drasing a second king given that one king was drawn first. 
```{r}
hands <- permutations(52,2 , v = deck)
first_card <-hands[,1]
second_card <- hands[,2]
sum(first_card %in% kings)
sum(first_card %in% kings & second_card %in% kings) / sum(first_card %in% kings)
```
What is the probability of a 21 in Black Jack
```{r}
aces <- paste("Ace" , suits)
facecard <- c("King" , "Queen" , "Jack" , "Ten")
facecard <- expand.grid(number = facecard , suit = suits)
facecard <- paste(facecard$number , facecard$suit)
hands <- combinations(52 , 2 , v = deck)
# probability of 21
mean(hands[,1] %in% aces & hands[,2]  %in% facecard)


```

